Push-Pull Signal-Function Functional Reactive Programming

### 2012-11-28 Docs/IFL2012/implementation.tex

The simplest example of the implementation of is the {\tt identity} combinator,
shown in Figure~\ref{figure:identity_implementation}. This signal function
simply passes all of its inputs along as outputs. The initialization function
simply passes along the received sample and outputs the initialized version of
the signal function. The initialized version of the input is similar, but is
self-referential. It outputs itself as its replacement. This is standard for
simple and routing combinators which are not reactive, and simply re-arrange,
discard, or combine signals and events.

@@

Combinators which take one or more signal functions as input must recursively
apply themselves, as is shown in the implementation of serial composition
(Figure~\ref{figure:serial_composition_implementation}). They must also
handle initialization, retaining the initialized signal functions and passing
them to the initialized version of the combinator.

@@

This combinator checks the outputs of the wrapped
signal function for an event occurrence from which an uninitialized signal
function is extracted. The switch combinator stores the full sample
for its input vector (which is identical to the input vector of the supplied
signal function) to initialize the new signal function. This also demands that
it add a wrapper to the new signal function which waits for the next sampling
interval and actuates the sample output at initialization as an output set
of changes to the signal. This has some performance implications, which are
discussed in Chapter~\ref{chapter:Evaluation_and_Comparisons}.

Most of the routing combinators are simple to implement. The only task is to add
remove, or replace routing constructors on signal updates and event occurrences.
Since these signal functions are stateless and primitive, they can simply
return themselves as their replacements. The {\tt swap} combinator is shown as
an example in Figure~\ref{figure:swap_implementation}

@@

The implementation strategy leaves room for optimizations. In particular, an
additional constructor for time-independent signal functions would allow
portions of a signal function to forgo evaluation during time steps unless they
had signal updates. Optimizations in the style of Yampa, observed by keeping
an updated AST for the signal function and pattern-matching on it when switching,
might further improve performance. In particular, collapsing nested or
serially-composed versions of the {\tt switchWait} step when switching would
remove at least some of the observed dependence of performance on sampling rate.
Nevertheless, this implementation performs quite well as it currently exists, as
we demonstrate in Chapter~\ref{chapter:Evaluation_and_Comparisons}.

@@

The evaluation interface is exported as shown in
Section~\ref{section:System_Design_and_Interface-Evaluator}.

@@

The {\tt update} action updates the current input signal delta, which will be
sampled at the next {\tt step} action. It has no immediately observable effect.
It simply updates the value of one signal in the input signal vector, a delta
of which is stored in the {\tt SFEvalState} record.

The {\tt step} action takes a time delta, and calls the signal continuation of
the stored signal function. It actuates the outputs using the stored handlers,
replaces the stored delta with an empty delta, and stores the resulting new
signal function in the state.

>>>>

The simplest example of the implementation of is the {\tt identity} combinator.
This signal function
simply passes all of its inputs along as outputs. The initialization function
simply passes along the received sample and outputs the initialized version of
the signal function. The initialized version of the input is similar, but is
self-referential. It outputs itself as its replacement. This is standard for
simple and routing combinators which are not reactive, and simply re-arrange,
discard, or combine signals and events.

@@

Combinators which take one or more signal functions as input must recursively
apply themselves, as is shown in the implementation of serial composition.
They must also
handle initialization, retaining the initialized signal functions and passing
them to the initialized version of the combinator.

@@

This combinator checks the outputs of the wrapped
signal function for an event occurrence from which an uninitialized signal
function is extracted. The switch combinator stores the full sample
for its input vector (which is identical to the input vector of the supplied
signal function) to initialize the new signal function. This also demands that
it add a wrapper to the new signal function which waits for the next sampling
interval and actuates the sample output at initialization as an output set
of changes to the signal. This has some performance implications, which are
discussed in Section~\ref{section:Discussion}.

Most of the routing combinators are simple to implement. The only task is to add
remove, or replace routing constructors on signal updates and event occurrences.
Since these signal functions are stateless and primitive, they can simply
return themselves as their replacements.

@@

The implementation strategy leaves room for optimizations. In particular, an
additional constructor for time-independent signal functions would allow
portions of a signal function to forgo evaluation during time steps unless they
had signal updates. Optimizations in the style of Yampa, observed by keeping
an updated AST for the signal function and pattern-matching on it when switching,
might further improve performance. In particular, collapsing nested or
serially-composed versions of the {\tt switchWait} step when switching would
remove at least some of the observed dependence of performance on sampling rate.
Nevertheless, this implementation performs quite well as it currently exists, as
we demonstrate in Section~\ref{section:Discussion}.

@@

@@

The {\tt update} action modifies the current input signal update, which will be
sampled at the next {\tt step} action. It has no immediately observable effect.
It simply updates the value of one signal in the input signal vector.

The {\tt step} action takes a time delta, and calls the signal continuation of
the stored signal function with the time delta and the stored signal update. It
actuates the outputs using the stored handlers, replaces the stored delta with
an empty delta, and stores the resulting new signal function in the state.
